---
pagetitle: "Reproducible Research"
author: "Maria Oros"
format:
  revealjs:
    css:
      - css/colors_and_type.css
      - css/uw-reveal.css
    transition: slide
    slide-number: true
    chalkboard: true
    code-fold: false
    code-line-numbers: true
execute:
  echo: true
---

::: {.uw-cover}
<div class="uw-topband"></div>

<h1 class="uw-cover-title">Reproducible Research</h1>
<div class="uw-mini-bar"></div>
<p class="uw-note" style="margin-top:0.6em;">Quarto and MyST Markdown</p>

::: {.uw-cover-meta}
<p class="name">Maria Oros</p>
:::

<img class="uw-crest" src="assets/brand/uw-crest-1color.svg" alt="">
:::

---

::: {.uw-slide}
<h2 class="uw-title">Agenda</h2>
<div class="uw-mini-bar"></div>

::: {.incremental}
- **The problem**: reproducibility crisis & core concepts
- **Literate programming**: code + narrative in one document
- **Quarto**: language-agnostic scientific publishing
- **MyST Markdown**: interactive books & mybinder
- **Reproducible environments**: renv, Docker, archiving
- **Best practices** & getting started
:::
:::

---

::: {.uw-section}
<div class="uw-sideband"></div>
<h2 class="uw-kicker">The Problem</h2>
<div class="uw-mini-bar"></div>
<p class="uw-section-sub">The reproducibility crisis, and the core concepts behind fixing it</p>
:::

---

::: {.uw-slide}
<h2 class="uw-title">What is Reproducible Research?</h2>
<div class="uw-mini-bar"></div>

The ability to recreate scientific findings using:

::: {.incremental}
- Original data
- Original code and analysis scripts
- Documentation of computational environment
- Clear methodology and workflow
:::

::: {.notes}
Reproducibility is fundamental to the scientific method
:::
:::

---

::: {.uw-slide}
<h2 class="uw-title">Why Reproducibility Matters</h2>
<div class="uw-mini-bar"></div>

::: {.columns}
::: {.column width="50%"}
**Credibility**

- Builds trust in findings
- Enables verification

**Efficiency**

- Saves time later
- Facilitates updates
:::

::: {.column width="50%"}
**Collaboration**

- Others can build on work
- Easier knowledge transfer

**Transparency**

- Clear methods
- Visible decisions
:::
:::
:::

---

::: {.uw-slide}
<h2 class="uw-title">The Reproducibility Crisis</h2>
<div class="uw-mini-bar"></div>

::: {.columns}
::: {.column width="55%"}
A 2016 *Nature* survey of 1,576 researchers found:

::: {.incremental}
- **>70%** failed to reproduce another scientist's experiments
- **>50%** failed to reproduce their *own* experiments
- Selective reporting & pressure to publish cited as top causes
:::
:::

::: {.column width="45%"}
::: {.callout-warning}
Irreproducible preclinical research alone is estimated to cost
**$28 billion/year** in the US.
:::
:::
:::

::: {.notes}
Baker, M. (2016). "1,500 scientists lift the lid on reproducibility." Nature.
:::
:::

---

::: {.uw-slide}
<h2 class="uw-title">Reproducible vs. Replicable</h2>
<div class="uw-mini-bar"></div>

| Term | Same data? | Same code? | Question answered |
|------|:---:|:---:|------|
| **Reproducible** | ✅ | ✅ | Do I get the same result from the same analysis? |
| **Replicable** | ❌ | ✅ | Does the finding hold with *new* data? |
| **Robust** | ✅ | ❌ | Does it hold with a *different* method? |
| **Generalizable** | ❌ | ❌ | Does it hold across data *and* methods? |

::: {.callout-note}
This deck focuses on **reproducibility** — the foundation the others build on.
:::
:::

---

::: {.uw-section}
<div class="uw-sideband"></div>
<h2 class="uw-kicker">Literate Programming</h2>
<div class="uw-mini-bar"></div>
<p class="uw-section-sub">Code and narrative living in one document</p>
:::

---

::: {.uw-slide}
<h2 class="uw-title">Literate Programming</h2>
<div class="uw-mini-bar"></div>

> "Instead of instructing a computer what to do, let us concentrate on
> explaining to humans what we want the computer to do." — Donald Knuth (1984)

::: {.columns}
::: {.column width="50%"}
**The idea**

- Code and narrative live in *one* document
- Prose explains the *why*; code shows the *how*
- Output is generated, never copy-pasted
:::

::: {.column width="50%"}
**Why it matters**

- No stale results from manual updates
- The document *is* the analysis
- Powers R Markdown, Quarto, MyST & Jupyter
:::
:::
:::

---

::: {.uw-section}
<div class="uw-sideband"></div>
<h2 class="uw-kicker">Quarto</h2>
<div class="uw-mini-bar"></div>
<p class="uw-section-sub">Language-agnostic scientific publishing</p>
:::

---

::: {.uw-slide}
<h2 class="uw-title">What is Quarto?</h2>
<div class="uw-mini-bar"></div>

> Next-generation scientific publishing system

- Successor to R Markdown
- **Language-agnostic**: R, Python, Julia, Observable
- Multiple output formats: HTML, PDF, Word, presentations
- Built-in support for citations and cross-references
- Developed by Posit
:::

---

::: {.uw-slide}
<h2 class="uw-title">Quarto Example: R</h2>
<div class="uw-mini-bar"></div>

```r
library(ggplot2)
library(dplyr)

# Load data
data <- data.frame(x = 1:10, y = (1:10)^2)

# Create plot
ggplot(data, aes(x = x, y = y)) +
  geom_line() + labs(title = "Quadratic Growth")
```
:::

---

::: {.uw-slide}
<h2 class="uw-title">Live R Code: Base Graphics</h2>
<div class="uw-mini-bar"></div>

```{r}
#| echo: true
#| output: true
#| fig-width: 8
#| fig-height: 4

# Generate data
x <- seq(0, 2*pi, length.out = 100)
y <- sin(x)

# Create plot
par(mar = c(4, 4, 2, 1))
plot(x, y, type = "l", lwd = 2, col = "blue",
     main = "Sine Wave",
     xlab = "x", ylab = "sin(x)")
grid()
```
:::

---

::: {.uw-slide}
<h2 class="uw-title">Quarto Example: ggplot2</h2>
<div class="uw-mini-bar"></div>

```{r}
#| echo: true
#| output: true
#| fig-width: 10
#| fig-height: 5

library(ggplot2)

# Create sample data
set.seed(42)
data <- data.frame(
  x = 1:10,
  y = rnorm(10, mean = 5, sd = 2)
)

# Create plot with ggplot2
ggplot(data, aes(x = x, y = y)) +
  geom_line(color = "steelblue", linewidth = 1.5) +
  geom_point(color = "darkred", size = 4) +
  theme_minimal(base_size = 14) +
  labs(title = "Random Data Visualization",
       x = "Index", y = "Value") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```
:::

---

::: {.uw-slide}
<h2 class="uw-title">Interactive Plot with plotly</h2>
<div class="uw-mini-bar"></div>

```{r}
#| echo: true
#| output: true
#| warning: false

library(plotly)

# Create sample data
data <- data.frame(
  x = 1:50,
  y = cumsum(rnorm(50))
)

# Create interactive plot
plot_ly(data, x = ~x, y = ~y, type = 'scatter', mode = 'lines+markers',
        line = list(color = 'rgb(55, 128, 191)', width = 2),
        marker = list(size = 6, color = 'rgb(219, 64, 82)')) %>%
  layout(title = "Interactive Random Walk",
         xaxis = list(title = "Step"),
         yaxis = list(title = "Value"))
```
:::

---

::: {.uw-slide}
<h2 class="uw-title">Quarto Key Features</h2>
<div class="uw-mini-bar"></div>

::: {.panel-tabset}
### Unified Syntax
Consistent markdown across all languages

### Rich Output
Interactive plots, tables, and widgets

### Publishing
Built-in support for websites, books, blogs

### Execution
Execute code in multiple engines

### Themes
Customizable templates and styling
:::
:::

---

::: {.uw-section}
<div class="uw-sideband"></div>
<h2 class="uw-kicker">MyST Markdown</h2>
<div class="uw-mini-bar"></div>
<p class="uw-section-sub">Interactive books, Jupyter, and mybinder</p>
:::

---

::: {.uw-slide}
<h2 class="uw-title">What is MyST Markdown?</h2>
<div class="uw-mini-bar"></div>

**M**arkedly **S**tructured **T**ext

- Extension of CommonMark Markdown
- Designed for scientific and technical documentation
- Part of the Jupyter ecosystem (Jupyter Book, Sphinx)
- Supports roles and directives from reStructuredText
- Excellent for creating interactive books
:::

---

::: {.uw-slide}
<h2 class="uw-title">MyST Markdown Syntax</h2>
<div class="uw-mini-bar"></div>

Here is example MyST syntax:

    # My Document

    :::{note}
    This is a special note block!
    :::

    Here's an equation: $E = mc^2$

    :::{admonition} Important
    :class: tip
    Remember to document your work!
    :::
:::

---

::: {.uw-slide}
<h2 class="uw-title">MyST Features</h2>
<div class="uw-mini-bar"></div>

::: {.columns}
::: {.column width="50%"}
**Directives**

    :::{note}
    Special blocks
    :::

**Roles**

    {kbd}`Ctrl+C`
    {file}`data.csv`

**Executable Notebooks**

    Jupyter notebooks as source
:::

::: {.column width="50%"}
**Cross-references**

    See {ref}`section-label`

**Math**

    $
    \int_0^1 x^2 dx = \frac{1}{3}
    $

**Interactive Computing**

    mybinder integration
:::
:::
:::

---

::: {.uw-slide}
<h2 class="uw-title">MyST + Jupyter Book Power</h2>
<div class="uw-mini-bar"></div>

::: {.columns}
::: {.column width="50%"}
**Built for Jupyter**

- Native notebook support
- Execute during build
- Cache outputs
- Interactive widgets
:::

::: {.column width="50%"}
**Publishing Features**

- Beautiful HTML books
- PDF via LaTeX
- Cross-references everywhere
- Automatic TOC
:::
:::

::: {.callout-note}
MyST excels at creating interactive documentation from Jupyter notebooks
:::
:::

---

::: {.uw-slide}
<h2 class="uw-title">What is mybinder?</h2>
<div class="uw-mini-bar"></div>

**mybinder.org**: Turn Git repositories into interactive notebooks

::: {.incremental}
- **Zero installation** for users
- **Reproducible environments** via configuration files
- **Share live code** with anyone
- **Free and open source**
:::

::: {.fragment}
Perfect combination: **MyST Markdown** + **mybinder** = Interactive, reproducible documentation
:::
:::

---

::: {.uw-slide}
<h2 class="uw-title">Why MyST + mybinder?</h2>
<div class="uw-mini-bar"></div>

::: {.columns}
::: {.column width="50%"}
### Traditional Documentation
- Static PDFs or HTML
- Code examples can't be run
- Users must set up environment
- Version mismatches
- Hard to experiment
:::

::: {.column width="50%"}
### MyST + mybinder
- Interactive notebooks
- **Click to run** in browser
- Environment pre-configured
- Always reproducible
- Easy experimentation
:::
:::

::: {.callout-tip}
Users can immediately interact with your analysis without any setup!
:::
:::

---

::: {.uw-slide}
<h2 class="uw-title">Real World Example: UW-Madison Survey</h2>
<div class="uw-mini-bar"></div>

[**Open Source Survey Results**](https://github.com/UW-Madison-DSI/open_source_survey_results_myst)

::: {.columns}
::: {.column width="60%"}
**Project Features:**

- Survey data analysis using MyST
- Jupyter Book documentation
- mybinder integration
- Fully reproducible workflow
- Interactive data exploration
:::

::: {.column width="40%"}
**Technologies:**

- MyST Markdown
- Jupyter notebooks
- Python analysis
- GitHub hosting
- Binder for compute
:::
:::
:::

---

::: {.uw-slide}
<h2 class="uw-title">UW-Madison Example Structure</h2>
<div class="uw-mini-bar"></div>

```bash
open_source_survey_results_myst/
├── _config.yml          # Jupyter Book config
├── _toc.yml            # Table of contents
├── requirements.txt     # Python dependencies
├── notebooks/          # Analysis notebooks
│   ├── demographics.ipynb
│   ├── usage_patterns.ipynb
│   └── visualizations.ipynb
├── data/               # Survey data
└── README.md           # With binder badge
```

::: {.fragment}
**Key insight**: Everything needed for reproduction in one repo!
:::
:::

---

::: {.uw-slide}
<h2 class="uw-title">mybinder Configuration Files</h2>
<div class="uw-mini-bar"></div>

::: {.panel-tabset}
### requirements.txt
```txt
pandas>=1.5.0
matplotlib>=3.5.0
seaborn>=0.12.0
jupyter-book>=0.13.0
myst-nb>=0.17.0
```

### environment.yml
```yaml
name: myproject
channels:
  - conda-forge
dependencies:
  - python=3.10
  - pandas
  - matplotlib
  - jupyter-book
```

### runtime.txt
```txt
r-2023-01-01
```
:::
:::

---

::: {.uw-slide}
<h2 class="uw-title">Adding mybinder to Your Project</h2>
<div class="uw-mini-bar"></div>

**1. Add a badge to README.md:**

```markdown
[![Binder](https://mybinder.org/badge_logo.svg)](
  https://mybinder.org/v2/gh/username/repo/main)
```

**2. Create requirements.txt or environment.yml**

**3. Push to GitHub**

**4. Click the badge - done!**

::: {.callout-tip}
mybinder automatically detects configuration files and builds your environment
:::
:::

---

::: {.uw-slide .smaller}
<h2 class="uw-title">MyST + mybinder Workflow</h2>
<div class="uw-mini-bar"></div>

```{mermaid}
flowchart TB
    A[Write MyST/Jupyter] --> B[Push to GitHub]
    B --> C[Jupyter Book Build]
    C --> D[Published Documentation]
    B --> E[mybinder Badge]
    E --> F[Click Badge]
    F --> G[Live Interactive Session]

    style D fill:#90EE90
    style G fill:#87CEEB
```

**Result**: Documentation that readers can **read** OR **execute**!
:::

---

::: {.uw-slide}
<h2 class="uw-title">Live Demo: Survey Analysis Example</h2>
<div class="uw-mini-bar"></div>

::: {.columns}
::: {.column width="50%"}
**In the Documentation:**

- Read analysis narrative
- View static plots
- See code and results
- Click "Launch Binder"
:::

::: {.column width="50%"}
**In mybinder:**

- Modify parameters
- Re-run analysis
- Create new plots
- Experiment freely
:::
:::

::: {.callout-important}
Same environment, guaranteed reproducibility!
:::
:::

---

::: {.uw-slide}
<h2 class="uw-title">MyST Advantages for Research</h2>
<div class="uw-mini-bar"></div>

1. **Jupyter-native**: Notebooks are first-class citizens
2. **Rich directives**: Admonitions, grids, cards, tabs
3. **Executable content**: Run code during build
4. **Cross-references**: Link figures, tables, sections
5. **mybinder integration**: One-click interactive sessions
6. **Version control friendly**: Plain text format
7. **Multiple outputs**: HTML books, PDFs, slides
:::

---

::: {.uw-slide}
<h2 class="uw-title">Example: MyST Document Header</h2>
<div class="uw-mini-bar"></div>

```yaml
---
jupytext:
  text_representation:
    extension: .md
    format_name: myst
kernelspec:
  display_name: Python 3
  language: python
  name: python3
---
```

This allows MyST to execute code blocks automatically!
:::

---

::: {.uw-slide}
<h2 class="uw-title">Example: Executable MyST Block</h2>
<div class="uw-mini-bar"></div>

```markdown
# Analysis Results

Here we analyze the survey data:

:::{code-cell} python
import pandas as pd
import matplotlib.pyplot as plt

# Load data
df = pd.read_csv('survey_data.csv')

# Create visualization
df['category'].value_counts().plot(kind='bar')
plt.title('Survey Responses by Category')
plt.show()
:::

The plot above shows clear trends in responses.
```

| Feature | Quarto | MyST Markdown |
|---------|--------|---------------|
| **Language Support** | Multi-language | Python-focused |
| **Output Formats** | HTML, PDF, PPT, Word | HTML, PDF |
| **Ecosystem** | Standalone | Jupyter/Sphinx |
| **Learning Curve** | Moderate | Steeper |
| **Best For** | Reports, websites, presentations | Books, documentation |
| **Interactive** | Yes (Observable, Shiny) | Yes (Jupyter widgets) |
:::

---

::: {.uw-slide}
<h2 class="uw-title">Comparison: Quarto vs MyST</h2>
<div class="uw-mini-bar"></div>

| Feature | Quarto | MyST Markdown |
|---------|--------|---------------|
| **Language Support** | Multi-language | Python-focused |
| **Output Formats** | HTML, PDF, PPT, Word | HTML, PDF |
| **Ecosystem** | Standalone | Jupyter/Sphinx |
| **Learning Curve** | Moderate | Steeper |
| **Best For** | Reports, websites, presentations | Books, documentation |
| **Interactive** | Yes (Observable, Shiny) | Yes (Jupyter widgets) |
| **mybinder** | Manual setup | Native integration |
:::

---

::: {.uw-slide .smaller}
<h2 class="uw-title">When to Use MyST?</h2>
<div class="uw-mini-bar"></div>

::: {.columns}
::: {.column width="50%"}
### Choose MyST When:

✅ Building interactive books

✅ Jupyter-centric workflow

✅ Need mybinder integration

✅ Python-heavy analysis

✅ Want Sphinx ecosystem

✅ Course materials online
:::

::: {.column width="50%"}
### Real Use Cases:

- **Research documentation** (UW-Madison example)
- **Course textbooks** with live code
- **Data analysis reports** with reproducible results
- **API documentation** with examples
- **Tutorial websites** with interactive exercises
:::
:::
:::

---

::: {.uw-slide}
<h2 class="uw-title">MyST Best Practices</h2>
<div class="uw-mini-bar"></div>

1. **Use Jupyter notebooks** as source files
2. **Add mybinder badge** for interactivity
3. **Document dependencies** clearly (requirements.txt)
4. **Cache outputs** for faster builds
5. **Use directives** for better formatting
6. **Cross-reference** figures and tables
7. **Version control** everything (including data when possible)
8. **Test mybinder link** before sharing
:::

---

::: {.uw-section}
<div class="uw-sideband"></div>
<h2 class="uw-kicker">Getting Started</h2>
<div class="uw-mini-bar"></div>
<p class="uw-section-sub">Installing and using Quarto and MyST</p>
:::

---

::: {.uw-slide}
<h2 class="uw-title">Getting Started: Quarto</h2>
<div class="uw-mini-bar"></div>

**Installation:**
```bash
# Download from quarto.org or use package manager
brew install quarto  # macOS
choco install quarto # Windows
```

**Create a document:**
```bash
quarto create-project myproject
cd myproject
quarto render document.qmd
```

**Preview:**
```bash
quarto preview document.qmd
```
:::

---

::: {.uw-slide}
<h2 class="uw-title">Getting Started: MyST</h2>
<div class="uw-mini-bar"></div>

**Installation:**
```bash
pip install jupyter-book myst-parser
```

**Create a book:**
```bash
jupyter-book create mybook/
cd mybook
jupyter-book build .
```

**Start with MyST-NB:**
```bash
pip install myst-nb
```
:::

---

::: {.uw-section}
<div class="uw-sideband"></div>
<h2 class="uw-kicker">Reproducible Environments</h2>
<div class="uw-mini-bar"></div>
<p class="uw-section-sub">renv, Docker, and archiving for the long run — plus best practices</p>
:::

---

::: {.uw-slide}
<h2 class="uw-title">Best Practices for Reproducibility</h2>
<div class="uw-mini-bar"></div>

1. **Version Control**: Use Git to track all changes
2. **Environment Management**:
   - Python: requirements.txt, conda env export
   - R: renv::snapshot(), renv.lock
3. **Clear Documentation**: Explain what, why, and how
4. **Modular Code**: Separate data, analysis, visualization
5. **Data Provenance**: Record sources and transformations
6. **Automated Workflows**: Use targets, make, drake
:::

---

::: {.uw-slide}
<h2 class="uw-title">Example: Complete Workflow</h2>
<div class="uw-mini-bar"></div>

```{mermaid}
flowchart LR
    A[Raw Data] --> B[Clean Data]
    B --> C[Analysis]
    C --> D[Visualizations]
    D --> E[Report]
    F[Code] --> C
    F --> D
    G[Parameters] --> C
    style E fill:#90EE90
    style A fill:#FFB6C1
```
:::

---

::: {.uw-slide .smaller}
<h2 class="uw-title">Practical Example: Data Analysis</h2>
<div class="uw-mini-bar"></div>

```{r}
#| echo: true
#| code-fold: true
#| output: true
#| warning: false
#| fig-width: 12
#| fig-height: 5

library(ggplot2)
library(dplyr)
library(tidyr)

# Set seed for reproducibility
set.seed(42)

# Create sample dataset
df <- data.frame(
  Method = rep(c('Traditional', 'Reproducible'), each = 50),
  Time_Saved = c(rnorm(50, mean = 5, sd = 2),
                 rnorm(50, mean = 15, sd = 3)),
  Errors = c(rpois(50, lambda = 5),
             rpois(50, lambda = 2))
)

# Create visualization
library(patchwork)

p1 <- ggplot(df, aes(x = Method, y = Time_Saved, fill = Method)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = c("#E69F00", "#56B4E9")) +
  labs(title = "Time Saved (hours/month)",
       x = NULL, y = "Hours") +
  theme_minimal() +
  theme(legend.position = "none")

p2 <- ggplot(df, aes(x = Method, y = Errors, fill = Method)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = c("#E69F00", "#56B4E9")) +
  labs(title = "Number of Errors",
       x = NULL, y = "Count") +
  theme_minimal() +
  theme(legend.position = "none")

p1 + p2
```
:::

---

::: {.uw-slide}
<h2 class="uw-title">Working with tidyverse</h2>
<div class="uw-mini-bar"></div>

```{r}
#| echo: true
#| output: true

library(dplyr)
library(tidyr)
library(knitr)

# Example data manipulation
mtcars %>%
  select(mpg, cyl, hp, wt) %>%
  group_by(cyl) %>%
  summarise(
    avg_mpg = mean(mpg),
    avg_hp = mean(hp),
    avg_wt = mean(wt),
    n = n()
  ) %>%
  kable(digits = 2, caption = "Summary Statistics by Cylinder")
```
:::

---

::: {.uw-slide}
<h2 class="uw-title">Statistical Modeling Example</h2>
<div class="uw-mini-bar"></div>

```{r}
#| echo: true
#| output: true

# Fit a linear model
model <- lm(mpg ~ wt + hp, data = mtcars)

# Display summary
summary(model)
```
:::

---

::: {.uw-slide}
<h2 class="uw-title">Model Diagnostics Plot</h2>
<div class="uw-mini-bar"></div>

```{r}
#| echo: true
#| output: true
#| fig-width: 10
#| fig-height: 6

# Diagnostic plots
par(mfrow = c(2, 2))
plot(model)
```
:::

---

::: {.uw-slide}
<h2 class="uw-title">Using renv for Reproducibility</h2>
<div class="uw-mini-bar"></div>

```{r}
#| eval: false
#| echo: true

# Initialize renv for your project
renv::init()

# Install packages
install.packages("ggplot2")
install.packages("dplyr")

# Take a snapshot of dependencies
renv::snapshot()

# Restore dependencies on another machine
renv::restore()
```

::: {.callout-tip}
## Best Practice
Always use renv to manage R package dependencies for reproducible research!
:::
:::

---

::: {.uw-slide .smaller}
<h2 class="uw-title">Containerization with Docker</h2>
<div class="uw-mini-bar"></div>

Capture the *entire* computational environment — OS, system libraries,
language, and packages — in a single portable image.

```dockerfile
FROM rocker/r-ver:4.3.0        # pinned R version

RUN R -e "install.packages('renv')"
COPY renv.lock renv.lock        # restore exact package versions
RUN R -e "renv::restore()"

COPY . /project
WORKDIR /project
CMD ["Rscript", "analysis.R"]
```

::: {.callout-tip}
`renv` pins packages; Docker pins *everything else*. Together they make an
analysis reproducible years later, on any machine.
:::
:::

---

::: {.uw-slide .smaller}
<h2 class="uw-title">Archiving & Citable Research</h2>
<div class="uw-mini-bar"></div>

::: {.columns}
::: {.column width="50%"}
**Make it permanent**

- **Zenodo** / **OSF** mint a **DOI** for your repo
- GitHub → Zenodo integration archives each release
- A DOI makes code & data *citable*
:::

::: {.column width="50%"}
**FAIR principles**

- **F**indable — persistent identifiers
- **A**ccessible — open, documented
- **I**nteroperable — standard formats
- **R**eusable — clear licensing
:::
:::

::: {.callout-note}
GitHub is for collaboration; an archive with a DOI is for *permanence*.
:::
:::

---

::: {.uw-slide}
<h2 class="uw-title">Creating Tables with gt</h2>
<div class="uw-mini-bar"></div>

```{r}
#| echo: true
#| output: true
#| warning: false

library(gt)
library(dplyr)

# Create a nice table
mtcars %>%
  head(10) %>%
  select(mpg, cyl, disp, hp, wt) %>%
  gt() %>%
  tab_header(
    title = "Motor Trend Car Road Tests",
    subtitle = "First 10 observations"
  ) %>%
  fmt_number(
    columns = c(mpg, disp, wt),
    decimals = 1
  ) %>%
  cols_label(
    mpg = "MPG",
    cyl = "Cylinders",
    disp = "Displacement",
    hp = "Horsepower",
    wt = "Weight"
  ) %>%
  tab_options(
    table.font.size = 12
  )
```
:::

---

::: {.uw-slide .smaller}
<h2 class="uw-title">Resources</h2>
<div class="uw-mini-bar"></div>

::: {.columns}
::: {.column width="50%"}
### Quarto
- [Official Docs](https://quarto.org)
- [Gallery](https://quarto.org/docs/gallery)
- [Awesome Quarto](https://github.com/mcanouil/awesome-quarto)

### MyST & Jupyter Book
- [MyST Parser](https://myst-parser.readthedocs.io)
- [Jupyter Book](https://jupyterbook.org)
- [MyST-NB](https://myst-nb.readthedocs.io)
- [mybinder](https://mybinder.org)
:::

::: {.column width="50%"}
### Examples
- [UW-Madison Survey](https://github.com/UW-Madison-DSI/open_source_survey_results_myst)
- [Executable Books Gallery](https://executablebooks.org/en/latest/gallery.html)

### General
- [The Turing Way](https://the-turing-way.netlify.app)
- [R for Data Science](https://r4ds.had.co.nz/)

### R Packages
- [tidyverse](https://www.tidyverse.org/)
- [renv](https://rstudio.github.io/renv/)
:::
:::
:::

---

::: {.uw-slide .smaller}
<h2 class="uw-title">Start Small: The Adoption Ladder</h2>
<div class="uw-mini-bar"></div>

You don't need everything at once — each rung is useful on its own and takes minutes.

::: {.incremental}
1. **Write in Quarto/R Markdown** — prose + one code chunk = literate programming *(zero setup)*
2. **`renv::snapshot()`** — one command pins your packages so it runs the same next year
3. **`requirements.txt` + Binder badge** — anyone runs your work in a browser, no install
4. **Docker or a Zenodo DOI** — only when you need exact environments or citable archives
:::

::: {.callout-tip}
The value compounds, but so does the effort — stop at whatever rung fits your needs.
:::
:::

---

::: {.uw-slide}
<h2 class="uw-title">Key Takeaways</h2>
<div class="uw-mini-bar"></div>

::: {.incremental}
1. **Reproducibility is the foundation** of credible science
2. **Literate programming** keeps code and narrative in sync
3. **Quarto & MyST** turn analyses into shareable, executable documents
4. **Pin your environment** — renv, Docker, `requirements.txt`
5. **mybinder** lets anyone run your work with zero setup
6. **Archive with a DOI** to make research permanent and citable
:::
:::

---

::: {.uw-cover .uw-closing}
<div class="uw-topband"></div>

<h1 class="uw-cover-title">Thank You!</h1>
<div class="uw-mini-bar"></div>
<p class="questions">Questions?</p>

::: {.uw-cover-meta}
<blockquote>
"Reproducible research is not just good practice—it's the foundation of credible science."
</blockquote>
<p class="venue">Try it out: check out the <a href="https://github.com/UW-Madison-DSI/open_source_survey_results_myst">UW-Madison example</a> and click the mybinder badge!</p>
:::

<img class="uw-crest" src="assets/brand/uw-crest-1color.svg" alt="">

::: {.notes}
Embrace reproducible research: Your future self will thank you!
:::
:::
