---
title: "Worked Example: A Reproducible Analysis"
subtitle: "Traditional vs. Reproducible workflows"
author: "UW-Madison DSI"
date: today
format:
  html:
    theme: cosmo
    css: styles.css
    toc: true
    toc-location: left
    code-fold: show
    fig-align: center
    fig-width: 8
    fig-height: 4.5
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
---

## Purpose

This document is a *small but complete* reproducible analysis. It reads the
dataset in `data/study_data.csv`, explores it, fits a model, and produces a
figure and a table — all from one source file. Re-render it and every number
below updates automatically.

```{r}
#| label: setup
library(readr)
library(dplyr)
library(ggplot2)
library(knitr)
```

## Load the data

We read the dataset with a path relative to the project root, so the analysis
runs the same on any machine.

```{r}
#| label: load
study <- read_csv(here::here("data", "study_data.csv"))
glimpse(study)
```

## Summarise

```{r}
#| label: summary
study |>
  group_by(method) |>
  summarise(
    n            = n(),
    mean_hours   = mean(hours_saved),
    mean_errors  = mean(errors),
    .groups = "drop"
  ) |>
  kable(digits = 2, caption = "Mean hours saved and errors by workflow")
```

## Visualise

```{r}
#| label: fig-hours
#| fig-cap: "Hours saved per month by workflow. The reproducible workflow shifts the whole distribution upward."
#| fig-width: 8
#| fig-height: 4.5
ggplot(study, aes(x = method, y = hours_saved, fill = method)) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.4) +
  scale_fill_manual(values = c("Traditional" = "#E69F00",
                               "Reproducible" = "#56B4E9")) +
  labs(title = "Time saved by workflow",
       x = NULL, y = "Hours saved / month") +
  theme_minimal(base_size = 13) +
  theme(legend.position = "none")
```

## Model

A simple linear model estimates the difference in hours saved between the two
workflows.

```{r}
#| label: model
model <- lm(hours_saved ~ method, data = study)
summary(model)
```

```{r}
#| label: tbl-effect
#| echo: false
effect <- coef(summary(model))["methodTraditional", "Estimate"]
```

The `Traditional` workflow is associated with about
**`r round(abs(effect), 1)` fewer hours saved per month** than the
`Reproducible` one — and because that number is computed from the data, it can
never drift out of sync with the table and figure above.

## Reproduce this analysis

```bash
# from the project root
quarto render analysis/report.qmd
```

That single command reruns everything and rebuilds `report.html`.
